---
title: "Identification of census variable vectors, Canadian Census 2011, 2016 and 2011"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

Codes below check census vector names from the `Cancensus` R package from census year 2011, 16, and 21.   
Target variables of interest (all proportion measures) and the corresponding denominators are:      
   - Visible minority   
   - Recent immigrant status    
   - First nation identify    
   - Education (highest degree as the secondary degree among those 15 over yaers old and among those between 25-65 years old, and having no secondary degree)      
   - Recent movers (moved past 1 year and 5 years)    
   - LICO (low income cutoff) - Prevalence measure, so there is no need to calculate denominator for this variable       
   - Property ownership, and those who pay >30% of income for housing, regardless of property status     
   - Age category, <15, 15-64, and >65. Census 2011 has different age categorization from the newer census, the numbers of age bing variables differ across census.       
  


```{r, message=FALSE, warning=FALSE }
# Libraries
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(tidyverse)
library(cancensus)
library(RPostgres)
library(rpostgis)
library(ggplot2)

rm(list=ls())

# cancensus
options(cancensus.cache_path = "cache/")


knitr::opts_chunk$set(echo = T,
                      results = "hide", 
                      warning = FALSE, 
                      message = FALSE)
                      

# function to check the value of census variables 
funcCheckCen <- function(yr, vec){
  get_census(dataset = yr, 
             regions = list("CSD" = "2434"),
             vectors = vec,
             level = "DA")
}

#list_census_datasets() %>% print(n = Inf)
#options(cancensus.cache_path = "cache/")
#CODES_TABLE %>% print(n = Inf)

list_census_datasets() %>% 
  filter(str_extract(description, "\\d+") %in% c("2021", "2016", "2011"))


df_vectorName <- data.frame(year = c("2021", "2016", "2011"))


```


#  Visible minority   
```{r}

# Show data vectors 
find_census_vectors("Visible", 
  dataset = "CA16", 
  type = "total", 
  query_type = "exact")

find_census_vectors("visible mino",
                    dataset = "CA16",
                    type = "total",
                    query_type = "semantic")

find_census_vectors("visible mino",
                    dataset = "CA11",
                    type = "total",
                    query_type = "semantic")


# Selected numerator and denominator
list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_4875", "v_CA21_4872"))

list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_3957", "v_CA16_3954"))

list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_457", "v_CA11N_460"))

# Display some numbers 
funcCheckCen("CA21", c("v_CA21_4875", "v_CA21_4872", "v_CA21_1")) %>%  
  select(-c(1:4), -c(8:10))
```

Save census vectors for denominators and numerators
```{r}

df_vectorName$visibleMin_count = c(
  "v_CA21_4875",
  "v_CA16_3957",
  "v_CA11N_460") 

df_vectorName$visibleMin_denom = c(
  "v_CA21_4872",
  "v_CA16_3954",
  "v_CA11N_457"
)
```
 


# Recent immigrants 
```{r}
find_census_vectors("Immigrant", 
                    dataset = "CA21", 
                    type = "total", 
                    query_type = "exact")


# Denominator nad numerator 
find_census_vectors("Immigrant", 
                    dataset = "CA16", 
                    type = "total", 
                    query_type = "exact")

find_census_vectors("Place of birth for the recent immigrant population in private households", 
                    dataset = "CA16", 
                    type = "total", 
                    query_type = "semantic")

# Denominator and numerator 
find_census_vectors("Immigrant", 
                    dataset = "CA11", 
                    type = "total", 
                    query_type = "exact")

find_census_vectors("recent immigrant", 
                    dataset = "CA11", 
                    type = "total", 
                    query_type = "semantic")


# display descriptors 
list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_4404","v_CA21_4635"))

list_census_vectors("CA16") %>% 
  filter(vector %in%  c("v_CA16_3405","v_CA16_3636"))

list_census_vectors("CA11") %>% 
  filter(vector %in%  c("v_CA11N_16","v_CA11N_265"))


# Display numbers to check 
funcCheckCen("CA21", c("v_CA21_4404","v_CA21_4635"))  %>%  select(-c(1:4), -c(8:10))
funcCheckCen("CA16", c("v_CA16_3405","v_CA16_3636")) %>%  select(-c(1:4), -c(8:10))
funcCheckCen("CA11", c("v_CA11N_16","v_CA11N_265"))%>%  select(-c(1:4), -c(8:10))


```

Save census vectors for denominators and numerators
```{r}
df_vectorName$recentImmig_count = c(
  "v_CA21_4635",
  "v_CA16_3636",
  "v_CA11N_265") 

df_vectorName$recentImmig_denom = c(
  "v_CA21_4404",
  "v_CA16_3405",
  "v_CA11N_16"
)
```
# First nation identity    
```{r}


find_census_vectors("Indigenous Identity", 
                    dataset = "CA21", 
                    type = "total", 
                    query_type = "semantic")

find_census_vectors("Indigenous Identity", 
                    dataset = "CA16", 
                    type = "total", 
                    query_type = "semantic")

find_census_vectors("Identity", 
                    dataset = "CA11", 
                    type = "total", 
                    query_type = "semantic")





list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_4201","v_CA21_4204"))  

list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_3855","v_CA16_3852"))  

list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_1351","v_CA11N_1354"))  



```
Save census vectors for denominators and numerators
```{r}
df_vectorName$indigi_count = c(
  "v_CA21_4204",
  "v_CA16_3855",
  "v_CA11N_1354") 

df_vectorName$indigi_denom = c(
  "v_CA21_4201",
  "v_CA16_3852",
  "v_CA11N_1351"
)



```
 

# Education - denominator (15 yrs old and 25-64 yrs old)
```{r}

# Denominators, >15 and 24-65
find_census_vectors("education total highest",
  dataset = "CA21",
  type = "total",
  query_type = "semantic") 


# Denominators, >15 and 24-65
find_census_vectors("education",
  dataset = "CA16",
  type = "total",
  query_type = "semantic") 


# Denominators, >15 and 24-65
find_census_vectors("education total",
                    dataset = "CA11",
                    type = "total",
                    query_type = "semantic") 


list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_5817", "v_CA21_5865"))

list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_5096", "v_CA16_5051"))

list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_1771", "v_CA11N_1801"))




```
Save census vectors for denominators and numerators
```{r}
# Record
df_vectorName$educ_HS_denom_15yr = c(
  "v_CA21_5817",
  "v_CA16_5051",
  "v_CA11N_1771") 

df_vectorName$educ_HS_denom_25_64yr = c(
  "v_CA21_5865",
  "v_CA16_5096",
  "v_CA11N_1801") 
```
 
# Education - two numerators (15 yrs old and 24-65 yrs old)
# Proportion of those whose highest degree is secondary (high school) degree, but not post-secondary nor non-high school degree 
Also get vectors for the count of non-high school finishers 
```{r}

find_census_vectors("high school secondary",
                    dataset = "CA21",
                    type = "total",
                    query_type = "semantic")

list_census_vectors("CA21") %>% 
  filter(grepl('v_CA21_58', vector)) %>% 
  filter(type == "Total") 

find_census_vectors("Total High (secondary) school diploma or equivalency certificate",
  dataset = "CA21",
  type = "total",
  query_type = "semantic")





## 2016 ------------------
find_census_vectors("education",
  dataset = "CA16",
  type = "total",
  query_type = "semantic")

list_census_vectors("CA16") %>% 
  filter(grepl('v_CA21_58', vector)) %>% 
  filter(type == "Total") 


find_census_vectors("education secondary",
                    dataset = "CA16",
                    type = "total",
                    query_type = "semantic")

find_census_vectors("no secondary",
                    dataset = "CA16",
                    type = "total",
                    query_type = "semantic")


## 2011 --------------------------------------

find_census_vectors("education secondary",
                    dataset = "CA11",
                    type = "total",
                    query_type = "semantic") 




# List variables ---- 
# Highest educ is high school 
list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_5823", "v_CA21_5871"))
# No high school educ 
list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_5820", "v_CA21_5868"))



list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_5057", "v_CA16_5102"))

list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_5144", "v_CA16_5330"))



list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_1777", "v_CA11N_1807"))

list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_1774", "v_CA11N_1804"))




```
Save census vectors for denominators and numerators

```{r}

# secondary degree holders 
df_vectorName$educ_HS_count_15yr = c(
  "v_CA21_5823",
  "v_CA16_5057",
  "v_CA11N_1777") 

df_vectorName$educ_HS_count_25_64yr = c(
  "v_CA21_5871",
  "v_CA16_5102",
  "v_CA11N_1807") 

# those with no degree 
df_vectorName$educ_no_HS_count_15yr = c(
  "v_CA21_5820",
  "v_CA16_5144",
  "v_CA11N_1774") 

df_vectorName$educ_no_HS_count_25_64yr = c(
  "v_CA21_5868",
  "v_CA16_5330",
  "v_CA11N_1804") 







```
 


# Recent movers, 1 and 5 yrs ago
```{r}
find_census_vectors("mover",
                    dataset = "CA21",
                    type = "total",
                    query_type = "semantic") 

find_census_vectors("mobility",
                    dataset = "CA16",
                    type = "total",
                    query_type = "semantic")

find_census_vectors("mobility",
                    dataset = "CA11",
                    type = "total",
                    query_type = "semantic")


# Numerator
list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_5751","v_CA21_5778")) 

# Numerator
list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_6725", "v_CA16_6698"))

# Numerator
list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_1723","v_CA11N_1750"))

# Denom 
list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_5745","v_CA21_5772"))
# Denom 
list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_6692", "v_CA16_6719"))
# Denom 
list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_1717","v_CA11N_1744"))


```
Save census vectors for denominators and numerators

```{r}
df_vectorName$move_1yr_count = c(
  "v_CA21_5751",
  "v_CA16_6698",
  "v_CA11N_1723") 

df_vectorName$move_5yr_count = c(
  "v_CA21_5778",
  "v_CA16_6725",
  "v_CA11N_1750") 



df_vectorName$move_1yr_denom = c(
  "v_CA21_5745",
  "v_CA16_6692",
  "v_CA11N_1717") 

df_vectorName$move_5yr_denom = c(
  "v_CA21_5772",
  "v_CA16_6719",
  "v_CA11N_1744") 
```
 

# Prevalance of LICO
```{r}
find_census_vectors("LICO",
                    dataset = "CA21",
                    type = "total",
                    query_type = "semantic")


find_census_vectors("LICO",
                    dataset = "CA16",
                    type = "total",
                    query_type = "semantic")


find_census_vectors("LICO",
                    dataset = "CA11",
                    type = "total",
                    query_type = "semantic")


list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_1085")) 

list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_2570"))

list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_2606"))

```
Save census vectors for denominators and numerators
```{r}

# Record
df_vectorName$LICO_prevalence = c(
  "v_CA21_1085",
  "v_CA16_2570",
  "v_CA11N_2606") 

```



# Property owners, and people paying >30% of income for housing regardless of property ownership 
```{r}

find_census_vectors(c("Owner"), 
                    dataset = "CA21", type = "total", query_type = "semantic") %>%  print(n=100)
find_census_vectors(c("Tenant"), 
                    dataset = "CA21", type = "total", query_type = "semantic") %>%  print(n=100)



find_census_vectors(c("Owner"), 
                    dataset = "CA16", type = "total", query_type = "semantic") %>%  print(n=100)
find_census_vectors(c("Tenant"), 
                    dataset = "CA16", type = "total", query_type = "semantic") %>%  print(n=100)



find_census_vectors(c("Owner"), 
                    dataset = "CA11", type = "total", query_type = "semantic") %>%  print(n=100)
find_census_vectors(c("Tenant"), 
                    dataset = "CA11", type = "total", query_type = "semantic") %>%  print(n=100)



# Denominator - total households with renting and owning , and income expenditure 
list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_4288", "v_CA21_4290", "v_CA21_4305", "v_CA21_4313"))

list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_4886", "v_CA16_4888", "v_CA16_4890", "v_CA16_4897"))

list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11N_2277", "v_CA11N_2281", "v_CA11N_2288", "v_CA11N_2279"))



funcCheckCen("CA21", c(
  "v_CA21_4288", # Owner and tenant households with household total income greater than zero, in non-farm, non-reserve private dwellings by shelter-cost-to-income ratio
  "v_CA21_4290", # Spending 30% or more of income on shelter costs
  "v_CA21_4305", # Total - Owner households in non-farm, non-reserve private dwellings
  "v_CA21_4313") # Total - Tenant households in non-farm, non-reserve private dwellings
  ) %>%  
select(-c(1:4), -c(8:10))


funcCheckCen("CA16", 
    c("v_CA16_4886", 
    "v_CA16_4888", 
    "v_CA16_4890", 
    "v_CA16_4897")
  ) %>%  
select(-c(1:4), -c(8:10))


funcCheckCen("CA11",  
    c("v_CA11N_2277", 
    "v_CA11N_2281", 
    "v_CA11N_2288", 
    "v_CA11N_2279")
  ) %>%  
select(-c(1:4), -c(8:10))


```
Save census vectors for denominators and numerators
```{r}

df_vectorName$property_denom = c(
  "v_CA21_4288",
  "v_CA16_4886",
  "v_CA11N_2277") 

df_vectorName$tenant_count = c(
  "v_CA21_4313",
  "v_CA16_4897",
  "v_CA11N_2288") 

df_vectorName$housingExpend_30_count = c(
  "v_CA21_4290",
  "v_CA16_4888",
  "v_CA11N_2279") 
```








# Age (1-15, 16-65, and >65)
2011 census data is 5 years bin,not like 2016 and 2021 that have a larger group like 0-14 range. It is possible to get denominator conunt from NHS (e.g. education age denominator like `v_CA21_5820`), but I suspect these denominators are from NHS. Thus, it might be safer to combine small age groups manually for 2021 census. 
```{r}

# 2021 and 2016
find_census_vectors("age", 
  dataset = "CA21", 
  type = "total", 
  query_type = "semantic") 


find_census_vectors("age", 
  dataset = "CA16", 
  type = "total", 
  query_type = "semantic") 



# 2011 
find_census_vectors("age", 
  dataset = "CA11", 
  type = "total", 
  query_type = "semantic") 

# Get more specifc age ranges, extract vector  
ageVec2011 <- find_census_vectors("Age Groups; Total population by age groups", 
  dataset = "CA11", 
  type = "total", 
  query_type = "semantic") %>%  print(n=200) %>% 
  select(vector)

# make tibble of census vectors into vector data 
ageVec2011 <- ageVec2011 %>%  pull()

# Display - do age category varaibles sum up to the "total population" column? 
df <- funcCheckCen("CA11", ageVec2011)  %>%  
  select(contains("_CA11F")) %>% 
  glimpse()




# Target age categories for 2021
list_census_vectors("CA21") %>% 
  filter(vector %in% c("v_CA21_8", "v_CA21_11", "v_CA21_68", "v_CA21_251"))

# Target age categories for 2016
list_census_vectors("CA16") %>% 
  filter(vector %in% c("v_CA16_1", "v_CA16_4", "v_CA16_61", "v_CA16_244"))

# Target age categories for 2011
list_census_vectors("CA11") %>% 
  filter(vector %in% c("v_CA11F_5", "v_CA11F_8", "v_CA11F_11", "v_CA11F_14", "v_CA11F_20"))

vecAge <- "v_CA11F_20:v_CA11F_23:v_CA11F_26:v_CA11F_29:v_CA11F_32:v_CA11F_35:v_CA11F_38:v_CA11F_41:v_CA11F_44:v_CA11F_47:v_CA11F_50:v_CA11F_53:v_CA11F_56:v_CA11F_59" 
vecAge <- str_split(vecAge, ":") %>% unlist()

list_census_vectors("CA11") %>% 
  filter(vector %in% vecAge) %>% 
  print(n = 30)

funcCheckCen("CA11", vecAge)  %>%  
  select(-c(1:4), -c(8:10)) %>% 
  select(Population, contains("years")) %>% 
  glimpse()



# Target age categories for 2021
funcCheckCen("CA21",  
  c("v_CA21_8", "v_CA21_11", "v_CA21_68", "v_CA21_251")) %>%  
  select(-c(1:4), -c(8:10)) %>% 
  glimpse()
# Target age categories for 2016
funcCheckCen("CA16", 
  c("v_CA16_1", "v_CA16_4", "v_CA16_61", "v_CA16_244"))  %>% 
  select(-c(1:4), -c(8:10)) %>% 
  glimpse()

# 2011 Age 0-14, more grouping than the other census 
funcCheckCen("CA11", 
  c("v_CA11F_5", "v_CA11F_8", "v_CA11F_11", "v_CA11F_14", "v_CA11F_20"))  %>% 
  select(-c(1:4), -c(8:10))  %>% 
  glimpse()
# 2011 Age 15-65, more grouping than the other census 
funcCheckCen("CA11", vecAge)  %>% 
  select(-c(1:4), -c(8:10)) %>% 
  glimpse()


```
Save census vectors for denominators and numerators
```{r}

df_vectorName$age_denom = c(
  "v_CA21_8",
  "v_CA16_1",
  "v_CA11F_5") 

df_vectorName$age_1_14_count = c(
  "v_CA21_11",
  "v_CA16_4",
  "v_CA11F_8:v_CA11F_11:v_CA11F_14") 

df_vectorName$age_15_64_count = c(
  "v_CA21_68",
  "v_CA16_61",
  "v_CA11F_20:v_CA11F_23:v_CA11F_26:v_CA11F_29:v_CA11F_32:v_CA11F_35:v_CA11F_38:v_CA11F_41:v_CA11F_44:v_CA11F_47:v_CA11F_50:v_CA11F_53:v_CA11F_56:v_CA11F_59") 

df_vectorName$age_64_over = c(
  "v_CA21_251",
  "v_CA16_244",
  NA) 
```


#### Transpose so that the columns will be census years
```{r}
rownames(df_vectorName) <- df_vectorName$year
dfVars <- df_vectorName %>% t %>% data.frame 
dfVars <- dfVars[row.names(dfVars) != "year", , drop = FALSE]

saveRDS(dfVars, "data/censMappedVars.rds")

# Following codes will extract the census vectors, along with polygons 
#b <- dfVars %>% pull(X2021)
#b <- setNames(b, rownames(a)) # Make it named vector so that variable names will be passed to extracted census vars  
#cen <- get_census(dataset='CA21', regions=list(CSD="5915022"),
#  vectors=b, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
#  api_key = Sys.getenv("CM_API_KEY"))
```





























