---
title: "CensusMap_Descriptive"
author: "H"

output:
  html_document:
    toc: true
    toc_float: true
---

Following codes create the map of socio-economic and demographic indicators from the data frame containing census vector of each year, which was created in the `censusVar_Mapping.Rmd` script. 

Also saves census data locally `data/city/listCen.rds`
And Saves DA shape files to remote DB as well, to be intersected to intervention lines. 


```{r setup, message=FALSE, warning=FALSE}

# Libraries
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(rgdal)
library(tidyverse)
library(cancensus)
library(ggmap)
library(stringr)
library(kableExtra)
library(htmlTable)
library(ggsankey)
library(grid)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(RPostgres)
library(rpostgis)
library(readxl)

rm(list=ls())

# Varlist for census 
source("function/varList.R")

# Canceuss 
#options(cancensus.cache_path = "cache/")
#set_cancensus_cache_path("cache/", install = TRUE)
options(cancensus.cache_path = "cache/.cancensus_cache")

bool_loadCensus = FALSE

# Plot global params 
theme_set(theme_classic())
theme_set(
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank(), 
    text = element_text(size=12), 
    plot.title = element_text(size=15),   
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
    )
)

# FUnctions for data preperation 
source("function/dataPrep.R")

# Load census vector names for each year 
dfCensVec <- readRDS("data/censMappedVars.rds")

```






Download census data based on the identified vector names, and create proporiton measures for each dissemination area. Rehspaed for facet map originally, but protting through looping. 
```{r}
funcGetCensVector <- function(df, dfVarDesc){
  # Calculate proportions
  df <- funcCensusIndicator(df)
  
  # Make it long if used for facet plot  
  dfMapLong <- df %>% 
    select(starts_with("i_")) %>% 
    pivot_longer(cols = starts_with("i_"), names_to = "x",values_to = "y")
  
  # Join sf with variable description
  dfMapLong <- dfMapLong %>% 
    left_join(dfVarDesc, by = c("x"="varName"))
  
  return(dfMapLong)
}


#csd_mtl <- c("2466127", "2466117", "2465005", "2466112","2466007", "2466023", "2466087", "2466097", "2466032", "2466107" , "2458227", "2458007")
```


### 2021 data 
```{r}
# Following codes will extract the census vectors, along with polygons 
b <- dfCensVec %>% pull(X2021)
b <- setNames(b, rownames(dfCensVec)) # Make it named vector so that variable names will be passed to extracted census vars  

mtl_2021 <- funcGetCenMontreal("CA21", b)  %>%
  st_transform(crs = 2959)

vic_2021 <- get_census(dataset='CA21', 
  regions=list(CSD = c("5917034")),
  vectors=b, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY")) %>%
  st_transform(crs = 3005)

van_2021 <- get_census(dataset='CA21', 
  regions=list(CSD = "5915022"),
  vectors=b, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY")) %>%
  st_transform(crs = 3153)

sk_2021 <- get_census(dataset='CA21', 
  regions=list(CSD = "4711066"),
  vectors=b, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY"))



```


### 2016 census 
```{r}
# Following codes will extract the census vectors, along with polygons 
b <- dfCensVec %>% pull(X2016)
b <- setNames(b, rownames(dfCensVec)) # Make it named vector so that variable names will be passed to extracted census vars  


mtl_2016 <- funcGetCenMontreal("CA16", b) %>%
  st_transform(crs = 2959)

vic_2016 <- get_census(dataset='CA16', 
  regions=list(CSD = c("5917034")),
  vectors=b, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY"))%>%
  st_transform(crs = 3005)

van_2016 <- get_census(dataset='CA16', 
  regions=list(CSD = "5915022"),
  vectors=b, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY")) %>%
  st_transform(crs = 3153)

sk_2016 <- get_census(dataset='CA16', 
  regions=list(CSD = "4711066"),
  vectors=b, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY"))




```

### 2011 census 
```{r}
b11 <- dfCensVec %>% pull(X2011)
b11 <- setNames(b11, rownames(dfCensVec)) # Make it named vector 

nameVec_age_1_14 <- paste("age_1_14_count",1:3, sep = "_")
nameVec_age_15_64 <- paste("age_15_64_count",1:14, sep = "_")

# Create extract individual census vector and get element names 
b11 <- c(b11, 
  str_split(b11[names(b11) == "age_1_14_count"], ":") %>%  unlist() %>% setNames(nameVec_age_1_14 ),
  str_split(b11[names(b11) == "age_15_64_count"], ":") %>%  unlist() %>% setNames(nameVec_age_15_64)
)
b11 <- b11[-c(grep(":", b11))] # remove array for age groups
b11 <- b11[!is.na(b11)] # remove NA for age 64 or greater, as it will be caclulated from other census vars 



### MTl 
#Mtl_2011<- get_census(dataset='CA11', 
#  regions=list(CSD = csd_mtl),
#  vectors=b11, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
#  api_key = Sys.getenv("CM_API_KEY"))

mtl_2011 <- funcGetCenMontreal("CA11", b11) %>%
  st_transform(crs = 2959)

mtl_2011$age_1_14_count <- mtl_2011 %>% 
  st_drop_geometry() %>%  
  select(starts_with("age_1_14")) %>%
  rowwise() %>% 
  rowSums(na.rm = FALSE) 

mtl_2011$age_15_64_count <- 
  mtl_2011 %>% 
  st_drop_geometry() %>%  
  select(starts_with("age_15_64")) %>% 
  rowwise() %>% 
  rowSums(na.rm = FALSE)

mtl_2011 <- mtl_2011 %>% 
  mutate(age_64_over = age_denom - (age_15_64_count + age_1_14_count)) 

### Victoria
vic_2011 <- get_census(dataset='CA11', 
  regions=list(CSD = "5917034"),
  vectors=b11, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY")) %>%
  st_transform(crs = 3005)

vic_2011$age_1_14_count <- vic_2011 %>% st_drop_geometry() %>%  select(starts_with("age_1_14")) %>% rowwise() %>% rowSums(na.rm = FALSE)
vic_2011$age_15_64_count <- vic_2011 %>% st_drop_geometry() %>%  select(starts_with("age_15_64")) %>% rowwise() %>% rowSums(na.rm = FALSE)

vic_2011 <- vic_2011 %>% mutate(age_64_over = age_denom - (age_15_64_count + age_1_14_count)) 



### Vancouver 
van_2011 <- get_census(dataset='CA11', 
  regions=list(CSD = "5915022"),
  vectors=b11, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY")) %>%
  st_transform(crs = 3153)

van_2011$age_1_14_count <- van_2011 %>% st_drop_geometry() %>%  select(starts_with("age_1_14")) %>% rowwise() %>% rowSums(na.rm = FALSE)
van_2011$age_15_64_count <- van_2011 %>% st_drop_geometry() %>%  select(starts_with("age_15_64")) %>% rowwise() %>% rowSums(na.rm = FALSE)

van_2011 <- van_2011 %>% mutate(age_64_over = age_denom - (age_15_64_count + age_1_14_count)) 


### SK
#sk_2011 <- get_census(dataset='CA11', 
#  regions=list(CSD = "4711066"),
#  vectors=b11, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
#  api_key = Sys.getenv("CM_API_KEY"))
#
#sk_2011$age_1_14_count <- sk_2011 %>% 
#  st_drop_geometry() %>%  
#  select(starts_with("age_1_14")) %>% 
#  rowwise() %>% rowSums(na.rm = FALSE)
#
#sk_2011$age_15_64_count <- sk_2011 %>% 
#  st_drop_geometry() %>%
#  select(starts_with("age_15_64")) %>% 
#  rowwise() %>% rowSums(na.rm = FALSE)
#
#sk_2011 <- sk_2011 %>% 
#  mutate(age_64_over = age_denom - (age_15_64_count + age_1_14_count)) 
```




### Get population rep points for Mtl , lon and lat are attached to each DA from canCensus 
```{r}
#length(unique(da$DAUID_ADIDU)) 6574
#length(unique(da$DARPLAT_ADLAT)) 6518
#length(unique(da$DARPLONG_ADLONG)) #6539

da <- read_csv("data/DA_center/2021_92_151_X.csv") %>% 
  filter(CMAPUID_RMRPIDU ==24462)
da_rep2021 <- da  %>%
  select(DAUID_ADIDU, DARPLONG_ADLONG, DARPLAT_ADLAT) %>% 
  group_by(DAUID_ADIDU, DARPLONG_ADLONG, DARPLAT_ADLAT) %>% 
  rename(dauid = DAUID_ADIDU, da_long = DARPLONG_ADLONG, da_lat = DARPLAT_ADLAT) %>% 
  summarise()  %>%  
  st_as_sf(coords = c("da_long", "da_lat"), crs = 4326, agr = "constant") %>%  
  st_transform(2959)
  #st_as_sf(coords = c("da_long", "da_lat"), crs = 4326, agr = "constant")


da <-read_excel("data/DA_center/2016_92_151_XBB.xlsx")
da_rep2016 <- da %>% 
  filter(`CMAuid/RMRidu` == 462) %>% 
  select(`DAuid/ADidu`, `DArplong/ADlong`, `DArplat/ADlat`) %>%  
  rename(dauid = `DAuid/ADidu`, da_long = `DArplong/ADlong`, da_lat = `DArplat/ADlat`) %>%  
  group_by(dauid, da_long, da_lat) %>%  
  summarise() %>%  
  st_as_sf(coords = c("da_long", "da_lat"), crs = 4326, agr = "constant") %>%  
  st_transform(2959)


da <-  read_excel("data/DA_center/2011_92-151_XBB_XLSX _columnName.xlsx")
da_rep2011 <- da %>% 
  filter(cmapuid  == 24462) %>% 
  select(DAuid, da_long, da_lat) %>%  
  rename(dauid = DAuid) %>% 
  group_by(dauid, da_long, da_lat) %>%  
  summarise() %>%  
  st_as_sf(coords = c("da_long", "da_lat"), crs = 4326, agr = "constant") %>%  
  st_transform(2959)

```



### Map population representative points   
```{r}
imp <- st_intersection(da_rep2021, mtl_2021)
ggplot(mtl_2011) +
  geom_sf(inherit.aes = FALSE, aes(fill = NULL), size =0.1, color= "black") + 
  geom_sf(data = imp, size =0.1) + 
  theme(legend.position = "bottom") + 
  theme(legend.text=element_text(size=9))




imp <- st_intersection(da_rep2011, mtl_2011)
ggplot(mtl_2011) +
  geom_sf(inherit.aes = FALSE, aes(fill = NULL), size =0.5, color= "black") + 
  geom_sf(data = imp, size =0.1) + 
  theme(legend.position = "bottom") + 
  theme(legend.text=element_text(size=9))




imp <- st_intersection(da_rep2016, mtl_2016)
ggplot(mtl_2016) +
  geom_sf(inherit.aes = FALSE, aes(fill = NULL), size =0.5, color= "black") + 
  geom_sf(data = imp, size =0.1) + 
  theme(legend.position = "bottom") + 
  theme(legend.text=element_text(size=9))

```




### Calculate distance Metro from each da in postgis
### Distance calc done by `distance_metroPoints.sql`
```{r}

metroDist <- st_read(conObesity, layer =  c("hiroshi", "da_metro_mtl_2021"), geometry_column ="geometry" ) # Euc distance 
mtl_2021 <- mtl_2021 %>% 
  mutate(GeoUID = as.numeric(GeoUID)) %>% 
  left_join(metroDist %>% st_drop_geometry(), by = join_by(GeoUID == dauid))

metroDist <- st_read(conObesity, layer =  c("hiroshi", "da_metro_mtl_2016"), geometry_column ="geometry" ) # Euc distance 
mtl_2016 <- mtl_2016 %>% 
  mutate(GeoUID = as.numeric(GeoUID)) %>% 
  left_join(metroDist %>% st_drop_geometry() %>% mutate(dauid = as.numeric(dauid)), by = join_by(GeoUID == dauid))

metroDist <- st_read(conObesity, layer =  c("hiroshi", "da_metro_mtl_2011"), geometry_column ="geometry" ) # Euc distance 
mtl_2011 <- mtl_2011 %>% 
  mutate(GeoUID = as.numeric(GeoUID)) %>% 
  left_join(metroDist %>% st_drop_geometry() %>% mutate(dauid = as.numeric(dauid)), by = join_by(GeoUID == dauid))



metroDist <- st_read(conObesity, layer =  c("hiroshi", "da_metro_van_2021"), geometry_column ="geometry" ) # Euc distance 
van_2021 <-van_2021 %>% 
  left_join(metroDist %>% st_drop_geometry() , by = join_by(GeoUID == GeoUID))

metroDist <- st_read(conObesity, layer =  c("hiroshi", "da_metro_van_2016"), geometry_column ="geometry" ) # Euc distance 
van_2016 <- van_2016 %>% 
  left_join(metroDist %>% st_drop_geometry(), by = join_by(GeoUID == GeoUID))

metroDist <- st_read(conObesity, layer =  c("hiroshi", "da_metro_van_2011"), geometry_column ="geometry" ) # Euc distance 
van_2011 <- van_2011 %>% 
  left_join(metroDist %>% st_drop_geometry() , by = join_by(GeoUID == GeoUID))


```


### Pop density  
```{r}
library(magrittr)
van_2011  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)
van_2016  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)
van_2021  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)
mtl_2011  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)
mtl_2016  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)
mtl_2021  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)
vic_2011  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)
vic_2016  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)
vic_2021  %<>%  mutate(popDensity_per1000 = Population/`Shape Area`/1000)

ggplot() + 
  geom_sf(data = van_2016, aes(fill= popDensity_per1000))

```


### Save data 
```{r save data}

### Representative points - Mtl only ----------------------
st_write( obj = da_rep2021, dsn = conObesity , layer = 'da_interact_mtl_2021_representative', delete_layer = TRUE)
st_write( obj = da_rep2016, dsn = conObesity , layer = 'da_interact_mtl_2016_representative', delete_layer = TRUE)
st_write( obj = da_rep2011, dsn = conObesity , layer = 'da_interact_mtl_2011_representative', delete_layer = TRUE)


### Montreal -----------------------------------------------------------------------------------------------
listCenAll <- list()
listCenAll[["2021"]] <- mtl_2021 %>% funcCensusIndicator 
listCenAll[["2016"]] <- mtl_2016 %>% funcCensusIndicator
listCenAll[["2011"]] <- mtl_2011 %>% funcCensusIndicator %>% select(-all_of(c(nameVec_age_15_64, nameVec_age_1_14)))
saveRDS(listCenAll, paste("data/Mtl/listCen.rds", sep = ""))

st_write( obj = listCenAll[["2021"]], dsn = conObesity , layer = 'da_interact_mtl_2021', delete_layer = TRUE)
st_write( obj = listCenAll[["2016"]], dsn = conObesity , layer = 'da_interact_mtl_2016', delete_layer = TRUE)
st_write( obj = listCenAll[["2011"]], dsn = conObesity , layer = 'da_interact_mtl_2011', delete_layer = TRUE)

### Victoria --------------------------------------------------------------------------------------------------
listCenAll <- list()
listCenAll[["2021"]] <- vic_2021 %>% funcCensusIndicator
listCenAll[["2016"]] <- vic_2016 %>% funcCensusIndicator
listCenAll[["2011"]] <- vic_2011 %>% funcCensusIndicator %>% select(-all_of(c(nameVec_age_15_64, nameVec_age_1_14)))
saveRDS(listCenAll, paste("data/Vic/listCen.rds", sep = ""))

st_write( obj = listCenAll[["2021"]], dsn = conObesity , layer = 'da_interact_vic_2021', delete_layer = TRUE)
st_write( obj = listCenAll[["2016"]], dsn = conObesity , layer = 'da_interact_vic_2016', delete_layer = TRUE)
st_write( obj = listCenAll[["2011"]], dsn = conObesity , layer = 'da_interact_vic_2011', delete_layer = TRUE)

### Vancouver -------------------------------------------------------------------------------------------------
listCenAll <- list()
listCenAll[["2021"]] <- van_2021 %>% funcCensusIndicator
listCenAll[["2016"]] <- van_2016 %>% funcCensusIndicator
listCenAll[["2011"]] <- van_2011  %>% funcCensusIndicator %>% select(-all_of(c(nameVec_age_15_64, nameVec_age_1_14)))
saveRDS(listCenAll, paste("data/Van/listCen.rds", sep = ""))

st_write( obj = listCenAll[["2021"]], dsn = conObesity , layer = 'da_interact_van_2021', delete_layer = TRUE)
st_write( obj = listCenAll[["2016"]], dsn = conObesity , layer = 'da_interact_van_2016', delete_layer = TRUE)
st_write( obj = listCenAll[["2011"]], dsn = conObesity , layer = 'da_interact_van_2011', delete_layer = TRUE)



```


Corr plots 
```{r fig.width = 10}


mtl_2021 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)

mtl_2016 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)

mtl_2011 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)



vic_2021 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)

vic_2011 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)

vic_2016 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)




van_2016 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)

vic_2011 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)

vic_2021 %>% 
  funcCensusIndicator  %>% 
  select(starts_with("i_")) %>% 
  st_drop_geometry() %>% 
  cor(use = "complete.obs") %>% 
  corrplot::corrplot.mixed(number.cex = 0.5, tl.cex = 0.5)



```


### Plot proportion maps for 2021 -  Scales vary across maps !-------------------------------------------
```{r fig.width=8, fig.height=8}
dfMapLong_mtl_2021 <- funcGetCensVector(mtl_2021, dfVarDesc)
dfMapLong_vic_2021 <- funcGetCensVector(vic_2021, dfVarDesc)
dfMapLong_van_2021 <- funcGetCensVector(van_2021, dfVarDesc)
#dfMapLong_sk_2021 <- funcGetCensVector(sk_2021, dfVarDesc)

dfMapLong_mtl_2016 <- funcGetCensVector(mtl_2016, dfVarDesc)
dfMapLong_vic_2016 <- funcGetCensVector(vic_2016, dfVarDesc)
dfMapLong_van_2016 <- funcGetCensVector(van_2016, dfVarDesc)
#dfMapLong_sk_2016 <- funcGetCensVector(sk_2016, dfVarDesc)

dfMapLong_mtl_2011 <- funcGetCensVector(mtl_2011, dfVarDesc)
dfMapLong_vic_2011 <- funcGetCensVector(vic_2011, dfVarDesc)
dfMapLong_van_2011 <- funcGetCensVector(van_2011, dfVarDesc)
#dfMapLong_sk_2011 <- funcGetCensVector(sk_2011, dfVarDesc)


# Generate plot with descriptive name 
pList_mtl_2021 <- list()
pList_vic_2021 <- list()
pList_van_2021 <- list()
#pList_sk_2021 <- list()

for(i in dfVarDesc$Description){
  pList_mtl_2021[[i]] <-  dfMapLong_mtl_2021 %>%
    filter(Description == i) %>% 
    ggplot() +
    geom_sf(inherit.aes = FALSE, aes(fill = y), size =0.5, color= "black") + 
    ggtitle(i) + 
    theme(legend.position = "bottom") + 
    theme(legend.text=element_text(size=9))
  
    pList_van_2021[[i]] <-  dfMapLong_van_2021 %>%
    filter(Description == i) %>% 
    ggplot() +
    geom_sf(inherit.aes = FALSE, aes(fill = y), size =0.5, color= "black") + 
    ggtitle(i) + 
    theme(legend.position = "bottom") + 
    theme(legend.text=element_text(size=9))
    
    pList_vic_2021[[i]] <-  dfMapLong_vic_2021 %>%
    filter(Description == i) %>% 
    ggplot() +
    geom_sf(inherit.aes = FALSE, aes(fill = y), size =0.5, color= "black") + 
    ggtitle(i) + 
    theme(legend.position = "bottom") + 
    theme(legend.text=element_text(size=9))
    
   # pList_sk_2021[[i]] <-  dfMapLong_sk_2021 %>%
  #  filter(Description == i) %>% 
   # ggplot() +
    #geom_sf(inherit.aes = FALSE, aes(fill = y), size =0.5, color= "black") + 
    #ggtitle(i) + 
    #theme(legend.position = "bottom") + 
    #theme(legend.text=element_text(size=9))
}
```

# MOntreal 
```{r}
pList_mtl_2021
```

# Vancouver 
```{r}
pList_van_2021
```

# VIctoria  
```{r}
pList_vic_2021
```

# Sask 
```{r}
#pList_sk_2021
```



