---
title: "intervention_analysis: Exploration of infrastrucure data by Year"
author: "H"
date: "2023-03-26"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}

# Libraries
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(rgdal)
library(tidyverse)
library(cancensus)
library(ggmap)
library(stringr)
library(kableExtra)
library(htmlTable)
library(ggsankey)
library(grid)
library(gridExtra)
library(magrittr)

rm(list=ls())

# Varlist for census 
source("function/varList.R")

# Canceuss 
options(cancensus.cache_path = "cache/")
#set_cancensus_cache_path("cache/", install = TRUE)

bool_loadCensus = FALSE


# Plot global params 
theme_set(theme_classic())
theme_set(
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank(), 
    text = element_text(size=12), 
    plot.title = element_text(size=15),   
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
    )
)

conObesity <- RPostgreSQL::dbConnect("PostgreSQL", host = "132.216.183.3",
  dbname = "obesity", user = "hiroshi", 
  password = Sys.getenv("DB3"))

```

### Load census and Dissemination area data - year 2021, from the CanCensus R library 
```{r, message=FALSE}

if(bool_loadCensus){
  listCen <- list()
  
  listCen[["Van"]] <- get_census(dataset='CA21', regions=list(CSD="5915022"),
  vectors=censusVarList, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY"))
  
  listCen[["Vic"]] <- get_census(dataset='CA21', regions=list(CSD="5917034"), vectors=censusVarList, level='DA', use_cache = FALSE, geo_format = "sf", quiet = TRUE, 
  api_key = Sys.getenv("CM_API_KEY"))
  
  # no need to run below anymore
  listCen[["Mtl"]] <- funcGetCenMontreal("CA21", censusVarList)

  saveRDS(listCen, "data/listCen_demo_2021.rds")
}else{
  listCen <- readRDS("data/listCen_demo_2021.rds")
}


```

Load intervention data, from Udit and Benoit 
THere are two maps for Mtl, based on two data sources
```{r, message=FALSE, warning=FALSE}
intList <- list() 

# Load latgest data, Sep 2023 
intList[["Vic"]] <- st_read("data/intervention_rawData/shp/updated_Sep2023/for_hiroshi/for_hiroshi/vic_bikelanes.shp") %>% st_transform(crs = 3005)

intList[["Van"]] <- st_read("data/intervention_rawData/shp/updated_Sep2023/for_hiroshi/for_hiroshi/van_bikelanes.shp") %>%st_transform(crs = 3153)

intList[["Mtl"]] <- st_read("data/intervention_rawData/shp/updated_Sep2023/mtl_bikelanes_no_mjr-road/mtl_bikelanes_no_mjr-road.shp") %>% st_transform(crs = 2959)

intList[["Mtl_other"]] <- st_read("data/intervention_rawData/shp/updated_Sep2023/mtl_bikelanes_no_mjr-road_other/mtl_bikelanes_no_mjr-road_other.shp") %>% st_transform(crs = 2959)


```






Load GTFS files for metro and bus stops 
```{r}
stops <- read.csv("data/GTFS_MTL/stops.txt")
metro_Mtl <- st_as_sf(stops, coords = c("stop_lon", "stop_lat"), crs = 4326) 
metro_Mtl <- metro_Mtl %>%  filter(location_type ==1) %>% st_transform(2959)

metro_Van <- read_sf("data/VAN_rapid-transit-stations/rapid-transit-stations.shp") %>% st_transform(3153)
```

 Data preperation - categorical varaibles for road types  
```{r, message=FALSE}

# New lane type 
intList[["Vic"]]$class_b <- factor(intList[["Vic"]]$facilities)
intList[["Mtl"]]$class_b <- factor(intList[["Mtl"]]$TypeVoie)
intList[["Mtl_other"]]$class_b <- factor(intList[["Mtl_other"]]$TypeVoie)
intList[["Van"]]$class_b <- factor(intList[["Van"]]$grnwy_n)

# Year name for Van 
intList[["Van"]]$yr <- intList[["Van"]]$year


l <- intList[["Vic"]] %>% st_length()/1000
intList[["Vic"]] %>%
  mutate(roadLen = l) %>%  
  group_by(class_b) %>%  
  st_drop_geometry() %>% 
  summarise(length = sum(roadLen)) 


l <- intList[["Van"]] %>% st_length() %>%  st_drop_geometry()
intList[["Van"]] %>%
  mutate(roadLen = l) %>%  
  group_by(class_b) %>%  
    st_drop_geometry() %>% 
  summarise(length = sum(roadLen)) 


l <- intList[["Mtl"]] %>% st_length()
intList[["Mtl"]] %>%
  mutate(roadLen = l) %>%  
  group_by(class_b) %>%  
    st_drop_geometry() %>% 
  summarise(length = sum(roadLen))  


l <- intList[["Mtl_other"]] %>% st_length()
intList[["Mtl_other"]] %>%
  mutate(roadLen = l) %>%  
  group_by(class_b) %>%  
    st_drop_geometry() %>% 
  summarise(length = sum(roadLen)) 

```

## Growth by year
#### Vic
```{r}
l <- intList[["Vic"]] %>% st_length()/1000
intList[["Vic"]] %>%
  mutate(roadLen = l) %>%  
  group_by(yr) %>%  
  st_drop_geometry() %>% 
  summarise(length = sum(roadLen)) 
```
#### Vancouver without stanley park
```{r}
map_noPark <- (listCen[["Van"]]) %>% filter(GeoUID != 59153586) %>% st_transform(3153)
intList[["Van"]] <- intList[["Van"]] %>% st_intersection(map_noPark)

l <- intList[["Van"]] %>% 
  st_length()/1000 %>%  
  st_drop_geometry()

intList[["Van"]] %>%
  mutate(roadLen = l) %>%  
  group_by(yr) %>%  
    st_drop_geometry() %>% 
  summarise(length = sum(roadLen)) 
```
#### Mtl, two versions of definitions from Benoit
```{r}
l <- intList[["Mtl"]] %>% st_length()/1000
intList[["Mtl"]] %>%
  mutate(roadLen = l) %>%  
  group_by(yr) %>%  
    st_drop_geometry() %>% 
  summarise(length = sum(roadLen))  


l <- intList[["Mtl_other"]] %>% st_length()/1000
intList[["Mtl_other"]] %>%
  mutate(roadLen = l) %>%  
  group_by(yr) %>%   
  st_drop_geometry() %>% 
  summarise(length = sum(roadLen)) 

```




### New data format, separate year layer to exsit_ format 
```{r} 



# Indicators for the existence of segment for each yr 
intList[["Vic"]] %<>%  
  mutate(exist_2011 = case_when(yr=="2011" ~ 1, TRUE ~ 0),
    exist_2016 = case_when(yr=="2016" ~ 1, TRUE ~ 0),
    exist_2021 = case_when(yr=="2021" ~ 1, TRUE ~ 0)) 


intList[["Van"]] %<>%  
  mutate(
    exist_2001 = case_when(yr=="2001" ~ 1, TRUE ~ 0),
    exist_2006 = case_when(yr=="2006" ~ 1, TRUE ~ 0),
    exist_2011 = case_when(yr=="2011" ~ 1, TRUE ~ 0),
    exist_2016 = case_when(yr=="2016" ~ 1, TRUE ~ 0),
    exist_2021 = case_when(yr=="2021" ~ 1, TRUE ~ 0)) 


intList[["Mtl"]] %<>%  
  mutate(
    exist_2001 = case_when(yr=="2001" ~ 1, TRUE ~ 0),
    exist_2006 = case_when(yr=="2006" ~ 1, TRUE ~ 0),
    exist_2011 = case_when(yr=="2011" ~ 1, TRUE ~ 0),
    exist_2016 = case_when(yr=="2016" ~ 1, TRUE ~ 0),
    exist_2021 = case_when(yr=="2022" ~ 1, TRUE ~ 0)) 


intList[["Mtl_other"]] %<>%  
  mutate(
    exist_2001 = case_when(yr=="2001" ~ 1, TRUE ~ 0),
    exist_2006 = case_when(yr=="2006" ~ 1, TRUE ~ 0),
    exist_2011 = case_when(yr=="2011" ~ 1, TRUE ~ 0),
    exist_2016 = case_when(yr=="2016" ~ 1, TRUE ~ 0),
    exist_2021 = case_when(yr=="2022" ~ 1, TRUE ~ 0)) 



intList[["Mtl"]]  %>%  group_by(class_b) %>% summarise(numSegments = n())
intList[["Mtl_other"]]  %>%  group_by(class_b) %>% summarise(numSegments = n())


```



#### Map of interventions 
```{r}


ggplot() + 
    geom_sf(data = intList[["Mtl"]] %>% filter(yr == 2022), aes(color = "Added in 2022")) + 
    geom_sf(data = intList[["Mtl"]] %>% filter(yr == 2016), aes(color = "Added in 2016")) + 
    geom_sf(data = intList[["Mtl"]] %>% filter(yr == 2011),  aes(color = "Existed in 2011")) +
    geom_sf(data = metro_Mtl, aes(color = "Metro")) +
  scale_color_manual(values = c("Added in 2022" = "blue", 
                                "Added in 2016" = "green",
                                "Existed in 2011" = "grey", 
                                "Metro" = "red"), 
                     name = "Year road cumulatively existed") 


ggplot() + 
    geom_sf(data = intList[["Mtl_other"]] %>% filter(yr == 2022), aes(color = "Added in 2022")) + 
    geom_sf(data = intList[["Mtl_other"]] %>% filter(yr == 2016), aes(color = "Added in 2016")) + 
    geom_sf(data = intList[["Mtl_other"]] %>% filter(yr == 2011),  aes(color = "Existed in 2011")) +
  scale_color_manual(values = c("Added in 2022" = "blue", 
                                "Added in 2016" = "green",
                                "Existed in 2011" = "grey" ), 
                     name = "Year road cumulatively existed") 



ggplot() + 
    geom_sf(data = intList[["Van"]] %>% filter(yr == 2021), aes(color = "Added in 2021")) + 
    geom_sf(data = intList[["Van"]] %>% filter(yr == 2016), aes(color = "Added in 2016")) + 
    geom_sf(data = intList[["Van"]] %>% filter(yr == 2011),  aes(color = "Existed in 2011")) +
    geom_sf(data = metro_Van, aes(color = "Station")) +
    scale_color_manual(values = c("Added in 2021" = "blue", 
                                "Added in 2016" = "green",
                                "Existed in 2011" = "grey",
                                "Station" = "red"), 
                     name = "Year road cumulatively existed") 


ggplot() + 
    geom_sf(data = intList[["Vic"]] %>% filter(yr == 2021), aes(color = "Added in 2022")) + 
    geom_sf(data = intList[["Vic"]] %>% filter(yr == 2016), aes(color = "Added in 2016")) + 
    geom_sf(data = intList[["Vic"]] %>% filter(yr == 2011),  aes(color = "Existed in 2011")) +
  scale_color_manual(values = c("Added in 2022" = "blue", 
                                "Added in 2016" = "green",
                                "Existed in 2011" = "grey" ), 
                     name = "Year road cumulatively existed") + 
  ggtitle("most of small segments will be removed (did not exist in photo image)")





```




 
 
 



```{r}
#### Finally, save the intervention data for the main analysis 
funcSaveGeom <- function(l){
  l$unique_id <- seq(1:nrow(l))
  l %<>% rename(geom = geometry)
}

intList <- lapply(intList, funcSaveGeom)
  
saveRDS(intList, "data/interventionList_annualLayer.rds")

st_write( obj = intList[["Van"]],  dsn = conObesity , layer = 'interact_intervention_annual_van', delete_layer = TRUE)

st_write( obj = intList[["Mtl"]], dsn = conObesity , layer = 'interact_intervention_annual_mtl', delete_layer = TRUE)

st_write( obj = intList[["Mtl_other"]], dsn = conObesity , layer = 'interact_intervention_annual_mtl_other', delete_layer = TRUE)


# Vic bike infras, removing short segments that did not exist phsyically
rd_2016 <- intList[["Vic"]] %>% filter(exist_2016 ==1 & !(unique_id %in% c(78,89,97,99,101,146,159)))
rd_2021 <- intList[["Vic"]] %>% filter(exist_2021 ==1 & !(unique_id %in% c(202,177,189,200,205,209,259,274)))
ggplot() + 
    geom_sf(data = intList[["Vic"]] %>% filter(exist_2016 ==1), linewidth=2,  color = "red", alpha = 0.9)  +
    geom_sf(data = rd_2016, linewidth=2,  color = "blue", alpha = 0.9)  
ggplot() + 
    geom_sf(data = intList[["Vic"]] %>% filter(exist_2021 ==1), linewidth=2,  color = "red", alpha = 0.9)  +
    geom_sf(data = rd_2021, linewidth=2,  color = "blue", alpha = 0.9)  
  

dim(intList[["Vic"]])
intList_trimmed <- intList[["Vic"]] %>% 
 filter(!(unique_id %in% c(78,89,97,99,101,146,159, 202,177,189,200,205,209,259,274)))
intList_trimmed

st_write( obj = intList_trimmed, dsn = conObesity , layer = 'interact_intervention_annual_vic', delete_layer = TRUE)
st_write( obj = intList[["Vic"]], dsn = conObesity , layer = 'interact_intervention_annual_vic_notTrimmed', delete_layer = TRUE)



# Save transit access 
st_write( obj = metro_Mtl, dsn = conObesity , layer = 'interact_metro_mtl', delete_layer = TRUE)
st_write( obj = metro_Van, dsn = conObesity , layer = 'interact_metro_van', delete_layer = TRUE)


```
