---
title: "Regression analysis: Vancouver minus S park "
author: "Hiroshi"

output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide 
---

```{r setup, include=FALSE}

# Libraries
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(rgdal)
library(tidyverse)
library(cancensus)
library(ggmap)
library(stringr)
library(kableExtra)
library(htmlTable)
library(ggsankey)
library(grid)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(keyring)
library(RPostgreSQL)
library(rpostgis)
library(magrittr)
library(CARBayes)

rm(list=ls())

# Varlist for census 
source("function/varList.R")

# Canceuss 
options(cancensus.cache_path = "cache/")
#set_cancensus_cache_path("cache/", install = TRUE)
bool_loadCensus = FALSE


# FUnctions for data preperation 
source("function/dataPrep.R")
source("function/plot_compile.R")


# Description of census variables 
dfVarDesc <- readRDS("~/projects/intervention_sdoh/data/varDescription.rds")

# Plot global params 
theme_set(theme_classic())

fileLoc <- "data/Van"
targetCMA <- "Van"
resultLoc <- "result/Van"

conObesity <- RPostgreSQL::dbConnect("PostgreSQL", 
  host = "132.216.183.3",
  dbname = "obesity", user = "hiroshi", 
  password = Sys.getenv("DB3"))


# Load
dfNetwork <- readRDS(paste(fileLoc,"/modelVars.rds", sep = ""))
dfNetwork$network_distance <- ifelse(dfNetwork$network_distance < dfNetwork$euc_distance | is.na(dfNetwork$network_distance), dfNetwork$euc_distance, dfNetwork$network_distance) 


# remove some variables that have too many zeros and thus block categorization by percentile 
# 2011, recent immig 

varMedianSplit <- "aboriginal"

```

# Data    

year of analysis: 2011, 2016, and 2021   
Infra definition follows Benoit's file excluding shallows (the two definitions provided earlier showed identical descriptive statistics, so  only one file is used here)


**Variable definitions**   
All variables are in proportion:   
  - `visMin`: visible minority  
  - `aboriginal`: First nation status **too many zeros, not used in the regression analysis that used tertile (not perssible to create percentile)**      
  - `educ_15`: proportion of residents whose highest education is high school diploma, thus excluding residents whose highest eudcation is post-secondary or no high-school diploma   
  - `educ_25-64`: same as above, but among those 25-64 years old **not added to the regression analysis**     
  - `move_5 yrs`: moved past 5 years   
  - `move_1 yrs`: moved past 1 years **not added to the regression analysis**    
  - `LICO`: Those under low-income threshold   
  - `rec_imm`: Recent immigrants (10 years) **Many zeros in Mtl**
  - `young`: <15 years old   
  - `old`: >64 years old     
  - `tenant_30 income`: those paying > 30% income for housing   **Not added to the regression model**   
  - `tenant`: tenant (not owner)  **Many zeros in Mtl**
  
```{r}
f <- formula(y ~ visMin + aboriginal +educ_15 + move_5yrs + LICO + rec_imm + young + old + tenant + min + popDensity_per1000)

# SES indicators 
modelCol <-c("visMin",  "educ_15", "move_5yrs", "LICO", "rec_imm", "young", "old", "tenant", "aboriginal")
```

  


# Descriptive 

### Proportion of DAs included in the analysis, after excluding DAs with missing missing (supresed) census indicators  
```{r, warning= FALSE}
cat("2011")
d <- dfNetwork %>% filter(yearPrevalent==2011)
d %>% st_drop_geometry() %>% complete.cases() %>% sum /nrow(d)
d %>% st_drop_geometry() %>% complete.cases() %>% sum


cat("2016")
d <- dfNetwork %>% filter(yearPrevalent==2016)
d %>% st_drop_geometry() %>% complete.cases() %>% sum /nrow(d)
d %>% st_drop_geometry() %>% complete.cases() %>% sum


cat("2021")
d <- dfNetwork %>% filter(yearPrevalent==2021)
d %>% st_drop_geometry() %>% complete.cases() %>% sum /nrow(d)
d %>% st_drop_geometry() %>% complete.cases() %>% sum

missingIdx <- complete.cases(dfNetwork %>% st_drop_geometry())
dfNetwork <- dfNetwork[missingIdx,]
```
### DAs in the study region with the outcomes, dichotomized access to bike infra and network distance 
**Year 2011 infra and census**  
**Only DA with non-missing variables are displayed, whited out areas are missing at least one census variable**
```{r, fig.width=7, fig.height=7}


ggplot(dfNetwork %>% filter(yearPrevalent == 2011)) + 
  geom_sf(aes(fill = prevalent)) + 
    theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank(), 
    text = element_text(size=12), 
    plot.title = element_text(size=15),   
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
    )

ggplot(dfNetwork %>% filter(yearPrevalent == 2011)) + 
  geom_sf(aes(fill=network_distance)) + 
    theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank(), 
    text = element_text(size=12), 
    plot.title = element_text(size=15),   
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12)
    )
```



### Histogram of census variables, 2011    
Large number of zero proportions in for educ, recent immigrants, and tenants for Van and Mtl
```{r, fig.width=12, fig.height=8, warning=FALSE}
pList <- list()
yr = 2011
for(i in gsub("i_", "", dfVarDesc$varName)){
  d <- dfNetwork %>% st_drop_geometry() %>% filter(yearPrevalent == yr)
  d <- data.frame(y=d$network_distance, x=d[, i] )
  pList[[i]] <- ggplot() +
    geom_histogram(data = d, aes(x=x), bins=40) + 
    ggtitle(i) 
}
do.call("grid.arrange", c(pList, ncol = 4))  

```

### Histogram of census variables, 2016    

```{r, fig.width=12, fig.height=8, warning=FALSE}

yr = 2016
for(i in gsub("i_", "", dfVarDesc$varName)){
  d <- dfNetwork %>% st_drop_geometry() %>% filter(yearPrevalent == yr)
  d <- data.frame(y=d$network_distance, x=d[, i] )
  pList[[i]] <- ggplot() +
    geom_histogram(data = d, aes(x=x), bins=40) + 
    ggtitle(i) 
}
do.call("grid.arrange", c(pList, ncol = 4))  

```

### Histogram of census variables, 2021    
```{r, fig.width=12, fig.height=8, warning=FALSE}

yr = 2021
for(i in gsub("i_", "", dfVarDesc$varName)){
  d <- dfNetwork %>% st_drop_geometry() %>% filter(yearPrevalent == yr)
  d <- data.frame(y=d$network_distance, x=d[, i] )
  pList[[i]] <- ggplot() +
    geom_histogram(data = d, aes(x=x), bins=40) + 
    ggtitle(i) 
}
do.call("grid.arrange", c(pList, ncol = 4))  


```  



### Corr of each variable  
```{r, fig.width=12, fig.height=8, warning=FALSE}
library("ggplot2") 
library("GGally") 
dfVarDesc$vatName_real <- gsub("i_", "", dfVarDesc$varName)

datCol  <- dfNetwork %>% 
  st_drop_geometry() %>% 
  dplyr::select(all_of(modelCol), yearPrevalent)

colnames(datCol) <- c("Visible min", "High Sch", "Moved", "Low income", "Recent immg", "Children",      "Seniors", "Tenant", "Aboriginal", "year")

datCol <- datCol  %>% 
 mutate(across(colnames(datCol)[-10], ~round(.x, 3)))

p <-ggpairs(datCol %>% filter(year=="2011") %>% dplyr::select(-year), 
    upper = list(continuous = wrap("cor", color = "black", size = 3)), 
    lower = list(continuous = wrap("points", color = "black", size = 0.05))) + theme_classic2(base_size = 7)
ggsave(plot = p, paste(resultLoc, "/Corr2011.jpg", sep = ""), width = 10, height = 7, unit = "in")

p <-ggpairs(datCol %>% filter(year=="2016") %>% dplyr::select(-year), 
    upper = list(continuous = wrap("cor", color = "black", size = 3)), 
    lower = list(continuous = wrap("points", color = "black", size = 0.05))) + theme_classic2(base_size = 7)
ggsave(plot = p, paste(resultLoc, "/Corr2016.jpg", sep = ""), width = 10, height = 7, unit = "in")

p <-ggpairs(datCol %>% filter(year=="2021") %>% dplyr::select(-year), 
    upper = list(continuous = wrap("cor", color = "black", size = 3)), 
    lower = list(continuous = wrap("points", color = "black", size = 0.05))) + theme_classic2(base_size = 7)
ggsave(plot = p, paste(resultLoc, "/Corr2021.jpg", sep = ""), width = 10, height = 7, unit = "in")

```  


### Outcome distribution, distance to bike infras in meter, 
Distance is **log transformed** (the non-log variables are highly skewed, highly likely to violate OLS assumptions). THus dependent varaiable for the linear model is natual log transformed distance, making the interpretation of regression parameter percent change, unlike Haude study.  
```{r, fig.width=4, fig.height=4}
# Distribution of continuous response
dfNetwork[dfNetwork$yearPrevalent == 2011, ]$network_distance %>% log() %>% hist(main = "Historgram of log distance")
dfNetwork[dfNetwork$yearPrevalent == 2016, ]$network_distance %>% log() %>% hist(main = "Historgram of log distance")
dfNetwork[dfNetwork$yearPrevalent == 2021, ]$network_distance %>% log() %>% hist(main = "Historgram of log distance")
```





### Scatter plot of distance (Y) and the proportion of census SES indicators (X), untransformd 
No linear trends, indicating the needs for transformation, polynomial modeling, or categorization (percentile bin) for valid modeling. I used categorical (tertile) and continuous without transformation (probably invalid) for modeling.   
#### 2011 , untransformed distance - terrible fit    
```{r, fig.width=12, fig.height=8, warning=FALSE}
pList <- list()
yr = 2011
for(i in gsub("i_", "", dfVarDesc$varName)){
  d <- dfNetwork %>% st_drop_geometry() %>% filter(yearPrevalent == yr)
  d <- data.frame(y=d$network_distance, x=d[, i] )
  pList[[i]] <- ggplot() +
    geom_point(data = d, aes(x=x, y=y)) + 
    ggtitle(i) 
}
do.call("grid.arrange", c(pList, ncol = 4))  
```

#### 2011 , Log transformed distance - still non-linear, but better  
```{r, fig.width=12, fig.height=8, warning=FALSE}
pList <- list()
yr = 2011
for(i in gsub("i_", "", dfVarDesc$varName)){
  d <- dfNetwork %>% st_drop_geometry() %>% filter(yearPrevalent == yr)
  d <- data.frame(y=log(d$network_distance), x=d[, i] )
  pList[[i]] <- ggplot() +
    geom_point(data = d, aes(x=x, y=y)) + 
    ggtitle(i) 
}
do.call("grid.arrange", c(pList, ncol = 4))  
```


# Regression analysis    

Prepare data 
```{r, fig.height=20, fig.width=3, warning=FALSE}

# log transformed Y and X, not used in this study 
library(rrapply)
is.word <- function(x) make.names(x) == x
insertLog <- function(x) substitute(log(x), list(x = x))
fLog <- rrapply(f, is.word, insertLog)

# Data to prep simulation  - make 1 percent increment for exposures
#modelDat <-  dfNetwork %>% 
##  dplyr::select(all_of(modelCol), min, popDensity_per1000, yearPrevalent, network_distance, prevalent) %>% 
#  mutate(across(modelCol, ~.x*10)) %>% 
#  mutate(across(modelCol, ~scale(.x, scale = F))) %>% 
#  st_drop_geometry()
```





### Spatial regression (hierarchical Bayesian) Model and non-spatial model, data prep  
Note that covariates are scaled to enhance mixing of markov chains for all models below 
```{r spatial }
# https://gis.stackexchange.com/questions/413159/how-to-assign-a-neighbour-status-to-unlinked-polygons
# Data to prep simulation  - make 1 percent increment for exposures
modelDat <-  dfNetwork %>% 
  dplyr::select(all_of(modelCol), min, popDensity_per1000, yearPrevalent, network_distance, prevalent) %>% 
  mutate(across(modelCol, ~.x*10)) %>% 
  group_by(yearPrevalent) %>% 
  mutate(across(modelCol, ~scale(.x, scale = F))) %>% 
  #mutate(min = scale(min, scale = T), popDensity_per1000 = scale((popDensity_per1000), scale = T))
  mutate(min = scale(log(min+0.000001), scale = T), popDensity_per1000 = scale(log(popDensity_per1000), scale = T))

```


#### Spatial weight matrix
**There is a a weird artifact in original SHP file, one pair of DA on the downdown area connecting across the water**
For valid results, weight matrices should have:     
- **symmetric entry**   
- **no isolated area (no zero in rowsum)**   
- **no isolated graph: rivers cannot divide the spatial window into several groups (shores) - not imporant for Van and VIC**     
```{R,  fig.height=12, fig.width=13, warning = FALSE, message = FALSE}  
library(spdep)
modelDat$y <- modelDat$network_distance 

wList <- list()
for(j in c("2011", "2016", "2021")){
    wList[[j]] <- funcGetW(modelDat, as.numeric(j))
}

plot(wList[[1]]$obj,  las=1, main='Connectivity by Shared Borders, isolated graphs by the water', sub = "not quite correct definition of neighbourhood",border = "gray")
plot(wList[[1]]$Wnb, coordinates(wList[[1]]$obj), pch = 0.1, cex = 0.1, points=FALSE, add=TRUE, col="black") #old (and invalid) spatial relationship

```



#### Run nonspatial (main analysis) and spatial (sub-analysis) OLS modelw,with Leroux CAR prior for spatial random effect 
##### Fitting with contionus scale of SES varaibles (likely to generate invalid results, results not shown)
```{r}
bool_runSpatial = FALSE
if(bool_runSpatial){
  fitListLmMonoCAR <- list( "2011"=list(), "2016"=list(), "2021" =list())
  fitListLmMonoNoCAR <- list( "2011"=list(), "2016"=list(), "2021" =list())
  modelDat$y <- modelDat$network_distance 
  
  for(j in c("2011", "2016", "2021")){
    for(i in modelCol){
      d <- modelDat %>% 
        filter(yearPrevalent == j) %>% 
        dplyr::select(all_of(i), y, min, popDensity_per1000) %>% 
        st_drop_geometry()
      
      fitListLmMonoCAR[[j]][[i]] <- S.CARleroux(formula=log(y)~., data=d, family="gaussian", W=wList[[j]]$W_mod, burnin=5000, n.sample=30000, thin=20, n.chains=3, n.cores=3)
      fitListLmMonoCAR[[j]][[i]]$family = "logGaussianCAR"
      
      fitListLmMonoNoCAR[[j]][[i]] = S.glm(formula=log(y)~., data=d, family="gaussian", burnin=5000, n.sample=30000, thin=20, n.chains=3, n.cores=3)
      fitListLmMonoNoCAR[[j]][[i]]$family = "logGaussianCAR"
      print(paste(j, i, sep = " ------------------------------------------- "))
    }
  }
  
  save.image()
  # Save data 
  saveRDS(fitListLmMonoCAR, paste(fileLoc,"/CAR.rds", sep = ""))
  saveRDS(fitListLmMonoNoCAR, paste(fileLoc,"/lm_mcmc.rds", sep = ""))
} else{
  #fitListLmMonoCAR <- readRDS(paste(fileLoc,"/CAR.rds", sep = ""))
  #fitListLmMonoNoCAR <- readRDS(paste(fileLoc,"/lm_mcmc.rds", sep = ""))

}
```

##### Fitting with categorical SES 
```{r Categorical OLS CAR}
#### Categorical SES -------------------------------------------------------------------------------------------------------------------

# Categorical Var
modelDatCat <-  modelDat %>% 
    group_by(yearPrevalent) %>% 
  dplyr::select(all_of(modelCol), min, popDensity_per1000, yearPrevalent, network_distance, prevalent) %>% 
  #mutate(across(all_of(modelCol) , ~ funcQuantile(.x, q=3))) %>% 
  mutate(across(all_of(modelCol[-which(modelCol == varMedianSplit)]) , ~ funcQuantile(.x, q=3))) %>% 
  #mutate(across(all_of(varMedianSplit) , ~ funcQuantile(.x, q=2))) %>% 
  data.frame() %>% 
  st_drop_geometry()   

# Custom subset of low proportion variable 
modelDatCat <-rbind(
  modelDatCat %>% 
    filter(yearPrevalent == "2011") %>% 
    mutate(across(all_of("aboriginal") , ~ funcQuantile(.x, q=6))),
  modelDatCat %>% 
    filter(yearPrevalent == "2016") %>% 
    mutate(across(all_of("aboriginal") , ~ funcQuantile(.x, q=2))),
  modelDatCat %>% 
    filter(yearPrevalent == "2021") %>% 
    mutate(across(all_of("aboriginal") , ~ funcQuantile(.x, q=3)))
)

# Outcome 
modelDatCat$y <- modelDatCat$network_distance 

bool_runSpatial_categorySES = FALSE
if(bool_runSpatial_categorySES){
  
  fitListLmMonoCAR_cat <- list( "2011"=list(), "2016"=list(), "2021" =list())
  fitListLmMonoNoCAR_cat <- list( "2011"=list(), "2016"=list(), "2021" =list())
  
  for(j in c("2011", "2016", "2021")){
    for(i in modelCol){
      d <- modelDatCat %>% 
        filter(yearPrevalent == j) %>% 
        dplyr::select(all_of(i), y, min, popDensity_per1000) 
        
      d[, i] <- droplevels(d[,i])
      
      fitListLmMonoCAR_cat[[j]][[i]] <- S.CARleroux(formula=log(y)~., data=d, family="gaussian", W=wList[[j]]$W_mod, burnin=5000, n.sample=100000, thin=10, n.chains=3, n.cores=3)
      fitListLmMonoCAR_cat[[j]][[i]]$family = "logGaussianCAR_cat"
      
      fitListLmMonoNoCAR_cat[[j]][[i]] = S.glm(formula=log(y)~., data=d, family="gaussian", burnin=5000, n.sample=30000, thin=10, n.chains=3, n.cores=3)
      fitListLmMonoNoCAR_cat[[j]][[i]]$family = "logGaussianCAR_cat"
      print(paste(j, i, sep = " ------------------------------------------- "))
    }
  }
  
  save.image()
  # Save data 
  saveRDS(fitListLmMonoCAR_cat, paste(fileLoc,"/CAR_cat.rds", sep = ""))
  saveRDS(fitListLmMonoNoCAR_cat, paste(fileLoc,"/lm_mcmc_cat.rds", sep = ""))
} else{
  fitListLmMonoCAR_cat <- readRDS(paste(fileLoc,"/CAR_cat.rds", sep = ""))
  fitListLmMonoNoCAR_cat <- readRDS(paste(fileLoc,"/lm_mcmc_cat.rds", sep = ""))
  
    # Give a label for family and variable type 
  fitListLmMonoCAR_cat[["2011"]] <- lapply(fitListLmMonoCAR_cat[["2011"]], function(x){ x$family = "logGaussianCAR_cat"; return(x)})
  fitListLmMonoCAR_cat[["2016"]] <- lapply(fitListLmMonoCAR_cat[["2016"]], function(x){ x$family = "logGaussianCAR_cat"; return(x)})
  fitListLmMonoCAR_cat[["2021"]] <- lapply(fitListLmMonoCAR_cat[["2021"]], function(x){ x$family = "logGaussianCAR_cat"; return(x)})
  
  fitListLmMonoNoCAR_cat[["2011"]] <- lapply(fitListLmMonoNoCAR_cat[["2011"]], function(x){ x$family = "logGaussianCAR_cat"; return(x)})
  fitListLmMonoNoCAR_cat[["2016"]] <- lapply(fitListLmMonoNoCAR_cat[["2016"]], function(x){ x$family = "logGaussianCAR_cat"; return(x)})
  fitListLmMonoNoCAR_cat[["2021"]] <- lapply(fitListLmMonoNoCAR_cat[["2021"]], function(x){ x$family = "logGaussianCAR_cat"; return(x)})
}

```

##### Fitting Logit spatial model- Leroux CAR model (NOT RUN)
```{r logit CAR}
bool_runSpatialBin = FALSE
if(bool_runSpatialBin){
  fitListBinMonoCAR <- list( "2011"=list(), "2016"=list(), "2021" =list())
  fitListBinMonoNoCAR<- list( "2011"=list(), "2016"=list(), "2021" =list())
  modelDatCat$y <- ifelse(modelDat$prevalent == "WithInfra", 1,)
  
  for(j in c("2011", "2016", "2021")){
    for(i in modelCol){
      d <- modelDatCat %>% 
        filter(yearPrevalent == j) %>% 
        dplyr::select(all_of(i), y, min, popDensity_per1000) %>% 
        st_drop_geometry()

      fitListBinMonoCAR[[j]][[i]] <- S.CARleroux(formula=y~., data=d, family="binomial", W=wList[[j]]$W_mod, burnin=15000, n.sample=50000, thin=1, n.chains=3, n.cores=3, trials = rep(1,nrow(d)))
      
      fitListBinMonoCAR[[j]][[i]]$family = "binomialCAR"
      
      fitListBinMonoNoCAR[[j]][[i]] = S.glm(formula=y~., data=d, family="binomial", burnin=15000, n.sample=50000, thin=10, n.chains=3, n.cores=3, trials = rep(1,nrow(d)))
      fitListBinMonoNoCAR[[j]][[i]]$family = "binomial"
      print(paste(j, i, sep = " ------------------------------------------- "))
    }
  }
  
  #fitListBinMonoNoCAR[["2011"]][["visMin"]]$samples$beta[[1]][, 3] %>%  plot()
  save.image()
  
  # Save data 
  saveRDS(fitListBinMonoCAR, paste(fileLoc,"/CARbin.rds", sep = ""))
  saveRDS(fitListBinMonoNoCAR, paste(fileLoc,"/bin_mcmc.rds", sep = ""))
}

```

### Check of model fit between spatial and non-spatial model, based on Watanabe-Akaike Information Criterion, lower value indicates better git 
```{r, warning=FALSE}
lapply(
  #fitListLmMonoCAR[["2011"]], function(x) x$modelfit[3]) %>% 
  #cbind(lapply(fitListLmMonoNoCAR[["2011"]], function(x) x$modelfit[3])) %>% 
  fitListLmMonoCAR_cat[["2011"]], function(x) x$modelfit[3]) %>% 
  cbind(lapply(fitListLmMonoNoCAR_cat[["2011"]], function(x) x$modelfit[3])) %>% 
  data.frame %>% 
  setNames(c("CAR_cat", "noCAR_cat")) 
```

  



### Table of results, OLS models that fits each SES variables and years separately.
**Interpretation** Since the distance variable (Y) was natural log transformed to mitigate terrible fit, the interpretation of the ses variables for the linear models are *percent* change. Specifically, the OLS coefficients indicate:   
**% changes in distance to bike infras by tertile level, relative to the distance at the lowest tertile level ( reference)** 

**Thus, higher values indicate worse access**.   


Results of logit model is not displayed here


### Forestplots for category SES vars
**Non-spatial models**
```{r, message = FALSE}
tb <- funcGetCoefTable(fitListLmMonoNoCAR_cat[["2011"]], categoryVar = T) %>% mutate(ID = 1:n()) %>% 
  full_join(funcGetCoefTable(fitListLmMonoNoCAR_cat[["2016"]], categoryVar = T) %>% mutate(ID = 1:n()), 
  by = "Var2") %>% 
  full_join(funcGetCoefTable(fitListLmMonoNoCAR_cat[["2021"]], categoryVar = T) %>% mutate(ID = 1:n()), 
  by = "Var2") %>% 
  arrange(ID) %>% 
  dplyr::select(-c(Var2, ID.x, Var.y, Var.x, ID.y, ID)) %>% 
  dplyr::select(Var, everything()) %>% 
  setNames(c("Var", rep(c("Mean (95% CI)"), 3))) 




a <- gsub("i_", "", dfVarDesc$varName)
b <- dfVarDesc$shortDesc
c <- tb$Var


repl <- tb$Var
for(i in seq_along(a)) repl <- gsub(a[i], b[i], repl, fixed = FALSE)
tb$Var <- repl

tb %>% 
  funcMakeTable(Title = "Categorical (tertile) SES for each model, Leroux CAR spatial model", rowlines = c(3, 6, 9, 12, 15, 18, 21, 24))

library(flextable)
library(officer)

write.table(tb, file = paste(resultLoc, "/coefTable.csv", sep = ""), sep = ",", quote = FALSE, row.names = T)

save_as_docx(
  "my table 1" = tb,  path = paste(resultLoc, "/coefTable.docx", sep = ""))  
```

**Spatial models**    
**Note that spatial model's fixed effects are shrunk towards zero after accouting for latent spatial effects captured by random effects, potentially due to spatial confounding**    
```{r}
tb <- funcGetCoefTable(fitListLmMonoCAR_cat[["2011"]], categoryVar = T) %>% mutate(ID = 1:n()) %>% 
  full_join(funcGetCoefTable(fitListLmMonoCAR_cat[["2016"]], categoryVar = T) %>% mutate(ID = 1:n()), 
  by = "Var2") %>% 
  full_join(funcGetCoefTable(fitListLmMonoCAR_cat[["2021"]], categoryVar = T) %>% mutate(ID = 1:n()), 
  by = "Var2") %>% 
  arrange(ID) %>% 
  dplyr::select(-c(Var2, ID.x, Var.y, Var.x, ID.y, ID)) %>% 
  dplyr::select(Var, everything()) %>% 
  setNames(c("Var", rep(c("Mean (95% CI)"), 3))) 

a <- gsub("i_", "", dfVarDesc$varName)
b <- dfVarDesc$shortDesc
c <- tb$Var

repl <- tb$Var
for(i in seq_along(a)) repl <- gsub(a[i], b[i], repl, fixed = FALSE)
tb$Var <- repl

tb %>% 
  funcMakeTable(Title = "Categorical (tertile) SES for each model, Leroux CAR spatial model", rowlines = c(3, 6, 9, 12, 15, 18, 21, 24))

library(flextable)
library(officer)

write.table(tb, file = paste(resultLoc, "/coefTable_spatial.csv", sep = ""), sep = ",", quote = FALSE, row.names = T)

save_as_docx(
  "my table 1" = tb,  path = paste(resultLoc, "/coefTable_spatial.docx", sep = ""))  

```



Codes to prepare data for graphing 
```{r, fig.height=15, fig.width=3, warning=FALSE}

funcGetCategoricalBetCarModel <- function(dList, year){
  lapply(dList[[as.character(year)]], funcGetCoef) %>%
    lapply(data.frame) %>% 
    lapply(function(x){ x %>% dplyr::filter(!rownames(x) %in% c("min", "popDensity_per1000", "nu2", "tau2", "rho", "(Intercept)" ))}) %>%
    bind_rows() %>% 
    mutate(yr = as.numeric(year)) %>% 
    rownames_to_column("Var") 
}


fitLmMonoCAR_cat <- rbind(
  fitListLmMonoCAR_cat %>% funcGetCategoricalBetCarModel(2011), 
  fitListLmMonoCAR_cat %>% funcGetCategoricalBetCarModel(2016), 
  fitListLmMonoCAR_cat %>% funcGetCategoricalBetCarModel(2021) 
) %>% 
  mutate(varName = str_sub(Var, 1, -2), varCat = str_sub(Var, -1, -1))

fitLmMonoNoCAR_cat <- rbind(
  fitListLmMonoNoCAR_cat %>% funcGetCategoricalBetCarModel(2011), 
  fitListLmMonoNoCAR_cat %>% funcGetCategoricalBetCarModel(2016), 
  fitListLmMonoNoCAR_cat %>% funcGetCategoricalBetCarModel(2021) 
) %>% 
  mutate(varName = str_sub(Var, 1, -2), varCat = str_sub(Var, -1, -1))
```





#### Plots for regression coefficients of tertile SES variables, non-spatial model 
```{r, fig.height=10, fig.width=9, warning=FALSE}
theme_barPlotText <-  theme(        
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12), 
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"), 
    title = element_text(size = 11)
)

# For now, remove aboriginal 
#modelCol <- modelCol[!modelCol %in% "aboriginal"]

pList <- list()

for(i in modelCol){
  dPlot <- fitLmMonoNoCAR_cat[fitLmMonoNoCAR_cat$varName == i, ] %>% 
    data.frame
  
  
    a <- gsub("i_", "", dfVarDesc$varName)
    b <- dfVarDesc$shortDesc
    repl <- i
    for(k in seq_along(a)) repl <- gsub(a[k], b[k], repl, fixed = FALSE)
    titleVarName <- repl
  
  pList[[i]] <- ggplot(data=dPlot, aes(y=Mean, x=factor(yr), color = varCat, shape = varCat)) + 
    geom_pointrange(aes(ymin = Lower, ymax = Upper), position = position_jitterdodge(jitter.width = 0.2), 
    dodge.width = 0.4,) + 
    ylim(c(-100, 180)) +
    geom_hline(yintercept=0, linetype="dashed", color = "black", linewidth=0.5) +
    ggtitle(titleVarName) + 
    xlab("Year") + ylab("Percent change") +
    scale_color_manual(
      values = c("red","blue"), 
      name = "Tertile of proportion     ", 
      labels = c("Middle tertile", "Highest tertitle")) +
    scale_shape_manual(
      values = c(17,19), 
      name = "Tertile of proportion     ", 
      labels = c("Middle tertile", "Highest tertitle")) + 
      theme_barPlotText
}

#do.call("grid.arrange", c(pList, ncol = 3))  
#ggarrange(plotlist = pList, ncol=4, nrow=2, common.legend = TRUE, legend="bottom")
ggarrange(plotlist = lapply(pList, "+", rremove("ylab")), ncol=4, nrow = 3, common.legend = TRUE, legend = "bottom") %>% 
  annotate_figure(left = textGrob("Percent change in distance relative to first tertile", rot = 90, vjust = 1, gp = gpar(cex = 1.1)))

ggsave(paste(resultLoc, "/OLS.jpg", sep = ""), height = 10, width=10, units = "in")

```



#### Plots for regression coefficients of tertile SES variables, spatial model 
**Again, note the huge attenuation of the estiamted mean of fixed effects, albeit smaller standard errors (higher precision)**
```{r, fig.height=10, fig.width=10, warning=FALSE}
pList <- list()

for(i in modelCol){
  dPlot <- fitLmMonoCAR_cat[fitLmMonoCAR_cat$varName == i, ] %>% 
    data.frame
  
    a <- gsub("i_", "", dfVarDesc$varName)
    b <- dfVarDesc$shortDesc
    repl <- i
    for(k in seq_along(a)) repl <- gsub(a[k], b[k], repl, fixed = FALSE)
    titleVarName <- repl
  
  pList[[i]] <- ggplot(data=dPlot, aes(y=Mean, x=factor(yr), color = varCat, shape = varCat)) + 
    geom_pointrange(aes(ymin = Lower, ymax = Upper), position = position_jitterdodge(jitter.width = 0.2), 
    dodge.width = 0.4,) + 
    ylim(c(-100, 180)) +
    geom_hline(yintercept=0, linetype="dashed", color = "black", linewidth=0.5) +
    ggtitle(titleVarName) + 
    xlab("Year") + ylab("Percent change") +
    scale_color_manual(
      values = c("red","blue"), 
      name = "Tertile of proportion      ", 
      labels = c("Middle", "Highest tertitle")) +
    scale_shape_manual(
      values = c(17,19), 
      name = "Tertile of proportion      ", 
      labels = c("Middle", "Highest tertitle")) + 
      theme_barPlotText
}

#do.call("grid.arrange", c(pList, ncol = 3))  
#ggarrange(plotlist = pList, ncol=4, nrow=2, common.legend = TRUE, legend="bottom")
ggarrange(plotlist = lapply(pList, "+", rremove("ylab")), ncol=4, nrow = 3, common.legend = TRUE, legend = "bottom") %>% 
  annotate_figure(left = textGrob("Percent change in distance relative to first tertile", rot = 90, vjust = 1, gp = gpar(cex = 1.1)))

ggsave(paste(resultLoc, "/OLS_spatial.jpg", sep = ""), height = 10, width=10, units = "in")
```








#### Show a subset of markov chain convergence, OLS with categorical varaibles
Unfold codes to see which SES variable and parameters were monitored for convergence 
```{r diagnosis_categoricalSPaital, fig.height42, fig.width=7}
fitListLmMonoCAR_cat[["2011"]][["visMin"]]$samples$beta[[2]][, 2] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["young"]]$samples$beta[[2]][, 2] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["tenant"]]$samples$beta[[2]][, 2] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["LICO"]]$samples$beta[[2]][, 2] %>%  plot()

fitListLmMonoCAR_cat[["2011"]][["visMin"]]$samples$tau2[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["young"]]$samples$tau2[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["tenant"]]$samples$tau2[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["LICO"]]$samples$tau2[[2]] %>%  plot()

fitListLmMonoCAR_cat[["2011"]][["visMin"]]$samples$nu[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["young"]]$samples$nu[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["tenant"]]$samples$nu[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["LICO"]]$samples$nu[[2]] %>%  plot()

fitListLmMonoCAR_cat[["2011"]][["visMin"]]$samples$rho[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["young"]]$samples$rho[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["tenant"]]$samples$rho[[2]] %>%  plot()
fitListLmMonoCAR_cat[["2011"]][["LICO"]]$samples$rho[[2]] %>%  plot()
```








# Test for the detection of effect sizes from Haude study, under mildly strong correlation of SES indicators
```{r}
# Check the prevalence of bike infra for intercept 
dfNetwork %>% 
  group_by(yearPrevalent) %>%  
  summarise(countDABikeInfra = sum(prevalent == "WithInfra") / n())
sum(dfNetwork$prevalent == "WithInfra") / nrow(dfNetwork)


# Data to prep simulation  
simDat <-  dfNetwork %>% 
  dplyr::select(all_of(modelCol), aboriginal, min, popDensity_per1000, yearPrevalent, network_distance, prevalent) %>% 
  dplyr::mutate(across(modelCol, ~.x*10))


funcDatGenerate <- function(s){
  # Prob no need for bootstrap for this simulation 
  # s <- sample_n(dfNetwork %>% st_drop_geometry(), N, replace = TRUE)

  # Outcome model - linear to logit, no interaction, no non-linear effect 
  xb <-
    4 + #baseline for 2016
    0*s$visMin + 
    #0*s$aboriginal +
    0*s$educ_15 +
    0*s$move_5yrs +
    0.3*s$LICO +
    -0.6*s$rec_imm +
    -1.6*s$young +
    -0.43*s$old +
    0*s$tenant + 
    #0*s$tenant_30income + 
    0.00005*s$min + 
    0.001*s$popDensity_per1000 
      
  s$xb <- xb # Logit for the binary outcomme 
  pi = 1/(1 + exp(-xb)) 
  s$y <- rbinom(n = length(xb), size = 1, prob = pi)

  # Response for liner model 
  xbLinear <-
    120 + #baseline for 2011
    0*s$visMin + 
    #0*s$aboriginal +
    0*s$educ_15 +
    0*s$move_5yrs +
    -80.1*s$LICO +
    320*s$rec_imm +
    600*s$young +
    127*s$old +
    0*s$tenant + 
    #0*s$tenant_30income + 
    0.00005*s$min + 
    0.001*s$popDensity_per1000 
  s$yl <- xbLinear
  return(s)
}

funcMySimRep <- function(s){
  s <- funcDatGenerate(s)
  # GLM 
  a <- glm(data = s, formula = f, family = "binomial" ) %>% summary
  
  s$y <- s$yl + rnorm(nrow(s),0,10)
  l <- lm(data = s, formula = f) %>% 
    summary
  return(list(a, l))
}

funcGetP <- function(simPerDf, var, family){
  simPerDf[[family]]$coefficients %>% 
    data.frame() %>%  
    rownames_to_column() %>% 
    filter(rowname == var) %>% 
    dplyr::select(starts_with("Pr"))  
}


# Simulation
yearTarget = 2021
simResults <- replicate(500, funcMySimRep(s= simDat %>% filter(yearPrevalent == yearTarget)), simplify = FALSE)


# Power for each exposure 
var="LICO"
a <- lapply(simResults, funcGetP, var, family = 1) %>%  unlist()   
sum(a < 0.05) /length(a) * 100
a <- lapply(simResults, funcGetP, var, family = 2) %>%  unlist()   
sum(a < 0.05) /length(a) * 100

rm(simDat)
```